\noappendix{List of Publications}%\label{sec:}

\sectionred*{Papers}

%\begin{itemize}

\incitem{2006_PLOS}{%
  E. Blanco, X. Messeguer, T.F. Smith and R. Guig\'{o}.\\
  ``Transcription factor map alignment of promoter regions.''\\
  \journal{PLoS Computational Biology}, 2: e49:403--416, 2006.}%

\vspace{0.5cm}

\incitem{2006_NAR}{%
  E. Blanco, D. Farr\'e, M. Alb\`a, X. Messeguer and R. Guig\'o.\\
  ``ABS: a database of Annotated regulatory Binding Sites\\ from orthologous promoters.''\\
  \journal{Nucleic Acids Research}, 34:D63--D67, 2006.}%

\vspace{0.5cm}

\incitem{2004_EMBO}{%
  S. Castellano, S.V. Novoselov, G.V. Kryukov, A. Lescure,\\ E. Blanco, A. Krol. V.N. Gladyshev and R. Guig\'o.\\
  ``Reconsidering the evolution of eukaryotic selenoproteins:\\ a novel non-mammalian family with scattered phylogenetic\\ distribution.''\\
  \journal{EMBO Reports}, 5:71--77, 2004.}%

\vspace{0.5cm}

\incitem{2003_PNAS}{%
   S. Beltran, E. Blanco, F. Serras, B. Perez-Villamil, R. Guig\'{o},\\ S. Artavanis-Tsakonas and M. Corominas.\\
  ``Transcriptional network controlled by the trithorax-group\\ gene ash2 in Drosophila melanogaster.''\\
  \journal{Proc. Nat. Acad. Sci.}, 100:3293--3298, 2003.}%

\vspace{0.5cm}

\incitem{2000_GenRes}{%
   G. Parra, E. Blanco and R. Guig\'{o}.\\
  ``Geneid in Drosophila.''\\
  \journal{Genome Research}, 10:511--515, 2000.}%

%\end{itemize}

\sectionred*{Book Chapters}

\incitem{2005_Baxevanis}{%
  E. Blanco and R. Guig\'{o}.\\
  ``Predictive Methods Using DNA Sequences.''\\
  In A. D. Baxevanis and B. F. Francis Ouellette, chief editors:\\
  \textbf{Bioinformatics: A Practical Guide to the Analysis of Genes\\ and Proteins, Third Edition.}\\
  John Wiley \& Sons Inc., New York, 2005. ISBN: 0--471--47878--4.}%

\vskip 1ex

\incitem{2002_Baxevanis}{%
  E. Blanco, G. Parra and R. Guig\'{o}.\\
  ``Using geneid to Identify Genes.''\\
  In A. D. Baxevanis and D. B. Davison, chief editors:\\
  \textbf{Current Protocols in Bioinformatics. Volume 1.}\\
  John Wiley \& Sons Inc., New York, 2002. ISBN: 0--471--25093--7.}%

\sectionred*{Posters}
%INCLUIR DESPUES EN FORMATO A4 LOS + RELEVANTES

\noindent E. Blanco, M. Pigantelli, X. Messeguer and R. Guig\'{o}.\\
  ``Deconstructing the position weight matrices to detect regulatory elements.''\\
  Global regulation of gene expression, Cold Spring Harbor, USA (2005)
\vskip 1.75ex

\noindent E. Blanco, X. Messeguer and R. Guig\'{o}.\\
  ``Novel computational methods to chracterize regulatory regions.''\\
  Genomic approaches to transcriptional regulation, Cold Spring Harbor, USA (2004)
\vskip 1.75ex

\noindent E. Blanco, X. Messeguer and R. Guig\'{o}.\\
  ``Alignment of promoter regions by mapping nucleotide sequences into arrays\\ of transcription factor binding motifs.''\\
  VII\textsuperscript{\textsl{th}} RECOMB, Berlin, Germany (2003)
\vskip 1.75ex

\noindent E. Blanco, G. Parra, S. Castellano, J.F. Abril, M. Burset,\\ X. Fustero, X. Messeguer and R. Guig\'{o}.\\
  ``Gene Prediction in the Post-Genomic Era.''\\
  IX\textsuperscript{\textsl{th}} ISMB, Copenhagen, Denmark (2001)
\vskip 1.75ex

\noindent J.F. Abril, M. Alb\`{a}, E. Blanco, M. Burset, F. C\^{a}mara, S. Castellano,\\ R. Castelo, O. Gonzalez, G. Parra and R. Guig\'{o}.\\
  ``Understanding the Eukaryotic Genome Sequence.''\\
  Inaugural Symposium of the Center for Genomic Regulation, Barcelona, Spain (2002)
\vskip 1.75ex

\noindent E. Blanco, G. Parra, S. Castellano, J.F. Abril, M. Burset, X. Fustero,\\ X. Messeguer and R. Guig\'{o}.\\
  ``Gene Prediction in the Post-Genomic Era.''\\
  IX\textsuperscript{\textsl{th}} ISMB, Copenhagen, Denmark (2001)
\vskip 1.75ex

\noindent J.F. Abril, E. Blanco, M. Burset, S. Castellano, X. Fustero, G. Parra and R. Guig\'{o}.\\
  ``Genome Informatics Research Laboratory: Main Research Topics.''\\
  I\textsuperscript{\textsl{st}} Jornadas de Bioinform\'{a}tica, Cartagena, Spain (2000)
\vskip 1.75ex


